---
title: "reshape tutorial"
author: "fg2399"
date: "2017/6/23"
output: github_document
---

#Reshape2 Tutoria

reshape2 is an R package that makes it easy to transform data between wide and long formats.

wide data has a column for each variable. what we see normally are usually wide data.

long-format data has a column for possible variable types and a column for the values of those variables. In reality, we need long-format data much commonly than wide-format data. For example, `ggplot2` requires long-format data(technically tidy data), `plyr` requires long-format data.

`reshape2` is based around two key functions: melt and cast.

##melt

melt takes wide-format data and melt it into long-format data.

```{r}
library(reshape2)
names(airquality) = tolower(names(airquality))
head(airquality)

#default melt
aql = melt(airquality, variable.name = "climate_variable", value.name = "climate_value")
head(aql)
tail(aql)
```

By default, `melt` has assumed that all columns with numeric values are variables with values. We can also add an ID variable that identify individual rows of data.

```{r}
#default + ID variable
aql = melt(airquality, id.vars = c("month", "day"),variable.name = "climate_variable", value.name = "climate_value")
head(aql)
```

##cast
In `reshape2` there are multiple `cast` functions. Since you will most commonly work with `data.frame` objects, we'll explore the `dcast` function.

Here, we need to tell `dcast` that `month` and `day` are the ID variables(we want a column for each) and that `variable` describes the measured variables. Since there is only one remaining column, `dcast` will figure out that it contains the values themselves. We could explicitly declare this with `value.var`

```{r}
aqw = dcast(aql, month + day ~ climate_variable)
head(aqw)

#you can also do some aggregation by using dcast

aqw = dcast(aql, month ~ climate_variable, fun.aggregate = mean)
head(aqw)


```


