---
title: "tidyr tutorial"
author: "fg2399"
date: "2017/6/27"
output: github_document
---

#tidyr tutorial

tidyr and dplyr provide fundamental functions for cleaning, processing, & manipulating data

##gather
`gather` function will reshape wide format to long format

It is pretty similar to `melt` function in `reshape`

```{r, message=FALSE}
library(tidyr)
library(reshape2)
head(airquality)

aql = gather(airquality, variable, value, -Month, -Day)
head(aql)

aql2 = melt(airquality, id.vars = c("Month","Day"), variable.name = "variable", value.name = "value")
head(aql2)
```

##spread
`spread` function will reshape long format into wide format

It is pretty similar to `dcast` function in `reshape2` 

```{r}
aqw = spread(aql, key = "variable", value = "value")
head(aqw)

aqw2 = dcast(aql, Month + Day ~ variable)
head(aqw2)
```

##unite

`unite` merge two varibales into one. It combines two varibales of a single observation into one variable.

```{r}
aql_united = unite(aql, col = date, Month, Day, sep = '.')
head(aql_united)
```


##seperate

`seperate` function will split a single variable into two.
It is similar to split the key into two variables, using a regular expression to describe the character that seperates them.
```{r}
aql_separate = separate(aql_united, col = date, into = c('Month','Day'), sep = '\\.')
head(aql_separate)
```




