---
title: "plyr tutorial"
author: "Fan Gong"
date: "2017/6/27"
output: github_document
---

#plyr tutorial

`plyr` is a data manipulating package which is especially for split-apply-combine.
In general, `plyr` is a set of tools that solves a common set of problems: you need to break a big problem down into manageable pieces, operate on each piece and then put all the pieces back together.

`plyr` builds on the build-in `apply` functions by giving you control over the input and output formats and keeping the syntax consistent across all variations. 

The basic format is two letters followed by `ply()`. The first letter refers to the format in and the second to the format out. 

The three main letters are: 

1. `d` = data frame 

2. `a` = array(include matrics) 

3. `l` = list

There are also some less common format options, here I will only talk about `_`, which means it throw away the output.

##split-apply-combine with plyr

###ddply
```{r}
library(plyr)
df = data.frame(year = rep(2000:2002, each = 3), count = round(runif(9, 0, 20)))
head(df)

#use ddply to calculate CV for every year
ddply(df, "year", function(x){
  mean = mean(x$count)
  sd = sd(x$count)
  cv = sd/mean
  return(data.frame(cv.count = cv))
})

```

###ldply
```{r}
#use ldply to calculate CV in every piece
df2 = data.frame(year = rep(2003:2005, each = 3), count = round(runif(9, 32, 50)))
df3 = data.frame(year = rep(2006:2008, each = 3), count = round(runif(9, 78, 90)))
l = list(period1 = df, period2 = df2, period3 = df3)

a = ldply(l, function(x){
  mean = mean(x$count)
  sd = sd(x$count)
  cv = sd/mean
  return(data.frame(cv.count = cv))
})

#also review the function of sapply
sapply(l, function(x){
  mean = mean(x$count)
  sd = sd(x$count)
  cv = sd/mean
  return(data.frame(cv.count = cv))
})

```

##Plotting with `plyr`

We can use `plyr` to plot data by throwing away the output with an underscore (_). This is a bit cleaner than a for loop since you don't have to subset the data manually.

```{r}
par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
d_ply(df, "year", transform, plot(count, main = unique(year), type = "o"))
mtext("count", side = 1, outer = TRUE, line = 1)
mtext("frequency", side = 2, outer = TRUE, line = 1)

```







